Modelling the gas kinematics of an atypical Lyman alpha emitting compact dwarf galaxy

### 2017-06-29 paper/paper.tex
abstract

\begin{abstract}
  The conditions in star-forming Compact Dwarf Galaxies (CDGs) 
  place them as the best systems to test models on primordial
galaxy formation and evolution.    
Here we report on one of such CDGs, \tol, which presents
a broad, single peaked, highly symmetric \lya\ line emission that had
evaded theoretical interpretation so far.  
In this paper we explain these features by two different physically
motivated kinematic models: 
an interstellar medium composed by outflowing clumps with 
random motions and an homogeneous gaseous sphere undergoing solid body
rotation.
It is the first time that an observed \lya\ spectrum can be explained
assuming either of these kinematic conditions.
The multiphase model requires a clump velocity dispersion of
at least \sigmaclump\ with outflows of \inftyclump\ at most, while the
bulk rotation velocity is constrained to be $348^{+75}_{-48}$ \kms.
This shows the importance of considering multiphase
and rotation kinematics among the possible conditions shaping the
\lya\ spectra.
However, we consider that the multiphase model has more chances to
agree with \tol's reality. 
The clump velocity dispersion implies a dynamical mass of at least $2\times
10^{9}$ M$_{\odot}$, ten times its baryonic mass, which can be
explained if \tol\ is hosted by a dark matter halo of at least
$10^{11}$ M$_{\odot}$; conditions similar to those of dwarf elliptical
galaxies. 
If future kinematic maps of \tol\ confirm the velocities found by the
multiphase model, it could provide a new perspective to study
evolutionary processes linking star-forming with quiescent phases in
primordial galaxy evolution. 
\end{abstract}

>>>>

\begin{abstract}

Star-forming Compact Dwarf Galaxies (CDGs) 
resemble the expected pristine conditions of the first galaxies in the
Universe and are the best systems to test models on primordial galaxy
formation and evolution. 
Here we report on one of such CDGs, \tol, which presents
a broad, single peaked, highly symmetric \lya\ emission line that had
evaded theoretical interpretation so far.  
In this paper we explain these features by two different physically
motivated kinematic models: 
an interstellar medium composed by outflowing clumps with 
random motions and an homogeneous gaseous sphere undergoing solid body
rotation.
It is the first time that an observed \lya\ spectrum can be explained
assuming either of these kinematic conditions.
The multiphase model requires a clump velocity dispersion of
at least \sigmaclump\ with outflows of \inftyclump\, while the
bulk rotation velocity is constrained to be $348^{+75}_{-48}$ \kms.
We argue that the multiphase model has more chances to be correct.
The clump velocity dispersion implies a dynamical mass of at least $2\times
10^{9}$ M$_{\odot}$, ten times its baryonic mass, which can be
explained if \tol\ is hosted by a dark matter halo of at least
$10^{11}$ M$_{\odot}$.
If future kinematic maps of \tol\ confirm the velocities suggested by the
multiphase model, it would provide additional support to expect this
condition to be present  in primordial \lya\ emitting galaxies.  
\end{abstract}
